Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-28 paper/results.tex
More
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@624 3fc257bb-8550-0410-81b3-a5c103da83f7

\subsection{R-MAT generation results}
Discuss R-MAT generation times.

@@

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview} 
and a multi-threaded
implementation
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.
In the multi-threaded MTGL~\cite{MTGL} implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

@@

\subsection{Maximally independent set results}
Discuss Luby maximally independent sets times.

@@

@@

To compare our MR-MPI implementation with a wider set of implementations,
we performed experiments MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
Cray XMT using two web graphs: {WebGraphA} with 13.2M
vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
The XMT and PBGL implementations are based on 
delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
over the entire edge list to advance a breadth-first search.  
We observe that the MR-MPI
implementation runs in less time than the Hadoop implementation, but requires
significantly more time than the XMT and PBGL implementations.  
The Bellman-Ford-style iterations are especially harmful to the MR-MPI and
Hadoop implementations for {WebGraphB}, which required 109 iterations to 
complete the search; 
execution times for this data set are show in Table~\ref{t:ssspB}.

@@

\begin{table}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI & 48 &  13,724 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{table}

>>>>

\subsection{R-MAT generation results}

In Figure~\ref{f:rmat}, we show the scalability of the R-MAT generation 
algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and RMAT-28.  For
RMAT-20 and RMAT-24, superlinear speed-up is shown.  This speed-up is due
to the decreased amount of file I/O needed with greater numbers of 
processors; with a larger total memory, a fixed-size problem requires
less I/O with more processors.  For RMAT-28, where most operation require
significant out-of-core operation, parallel efficiency ranges from 52\% to
97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
the original algorithm (Figure~\ref{fig:rmat}) showed approximately 
10\% reduction in execution time for the enhanced algorithm.

KDD NOTE:  Add XMT or PBGL results here.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
\label{f:rmat}
\end{figure}

@@

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
(Figure~\ref{fig:pr2}) compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview} 
and a multi-threaded
implementation
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.
In the multi-threaded MTGL~\cite{MTGL} implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

@@

\subsection{Maximally independent set results}

The execution times for the maximally independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
generation results, superlinear speed-up of the algorithm occurs for RMAT-20
and RMAT-24, as more of the graph fits into processor memory and less file
I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
operations. In this case, parallel efficiency is nearly perfect going from 
8 to 32 processors; some loss of efficiency is seen at 64 processors,
where the parallel efficiency compared to the 8-processor case is ??\%.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

@@

To compare our MR-MPI implementation with a wider set of implementations,
we performed experiments MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
Cray XMT using two web graphs: {WebGraphA} with 13.2M
vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
The XMT and PBGL implementations are based on 
delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
over the entire edge list to advance a breadth-first search.  
We observe that the MR-MPI
implementation runs in less time than the Hadoop implementation,but requires
significantly more time than the XMT and PBGL implementations.  
In experiments with {WebGraphB},
the benefit of the enhanced algorithm (Figure~\ref{fig:sssp2}) is clearly
shown, with a 40\% reduction in execution time compared to the original
SSSP algorith (Figure~\ref{fig:sssp}).
But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
Hadoop implementations for {WebGraphB}, which required 110 iterations to 
complete; execution times for this data set are show in Table~\ref{t:ssspB}.

@@

\begin{table}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{table}
